---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    keep_md: true
self_contained: false

---

```{r}
#load packages
library(DOS2)
library(optmatch)
library(RItools)
library(ggplot2)
library(dplyr)
source('utility.R')
library(knitr)
library(sensitivitymv)
fludata <- read.csv("fludata.txt", sep="")

```

## Check covariate balance

```{r include=FALSE}
#check covariate balance for everyone
test = plot(xBalance(receive ~ assign + age + copd + dm + heartd + race + renal + sex + liverd - 1, data=fludata))

png(filename="matching/balance_pre_any_matching.png")
print(test)
dev.off()
share_received = mean(fludata$receive) #share that was encouraged
n_receive = count(fludata, receive) #number of treated and untreated units among those not encouraged
```


## TO-DO: Other checks of how covariates are distributed for treated and untreated

##define some functions

```{r}
plot_discrete = function(data, treat, var_of_interest) {
  treat = sym(treat)
  var_of_interest = sym(var_of_interest)
  
  freqs = data %>% 
    mutate(!!var_of_interest := as.factor(!!var_of_interest)) %>% 
    group_by(!!treat, !!var_of_interest) %>%
    summarize(n = n()) %>%
    mutate(pct = n/sum(n))
  
  ggplot(freqs, aes(x = !!var_of_interest, y = pct, fill = !!treat)) +
     geom_bar(stat = "identity", position = "dodge")  +
      ggtitle(paste0("Relative Frequency of ", var_of_interest, " by treatment status")) +
      labs(fill = "Treatment Status")
}

plot_rel_freq <- function(var, treat, data) {
  
  var = sym(var)
  treat = sym(treat)
  
  ggplot(data, aes(x = !!var, fill = !!treat)) +
    geom_density(alpha = 0.5) +
    ggtitle(paste0("Density of ", var, " by treatment status")) +
    labs(fill = "Treatment Status")
}

```

# Compute prop scores and covariate distances

```{r}
#Compute propensity score 
fludata$prop <- glm(receive ~ assign + age + copd + dm + heartd + race + renal + sex + liverd, family=binomial, data=fludata)$fitted.values

#Plot propensity score by treatment status
ggplot(fludata, aes(x = prop, fill = as.factor(receive))) +
  geom_density(alpha = 0.5) +
  ggtitle(paste0("Density of propensity score by treatment status")) +
  labs(fill = "Treatment Status")
ggsave("matching/propensity_score_by_treatment.png")

#Plot propensity score by COPD status
ggplot(fludata, aes(x = prop, fill = as.factor(copd))) +
  geom_density(alpha = 0.5) +
  ggtitle(paste0("Density of propensity score by COPD")) +
  labs(fill = "COPD")
ggsave("matching/propensity_score_by_copd.png")

#Compute covariate distances
fludata= fludata%>% mutate(z = receive) ##otherwise the summary function does not work. fix this later, cos Z it's usually reserved for the instrument, not for the instrumented variable.
fludata= fludata%>% mutate(id = row_number())

```

##Generate 1:1 dataset
```{r}
#Baseline 1:1 dataset
dist1 <- match_on(z ~ assign + age + copd + dm + heartd + race + renal + sex + liverd, method = "mahalanobis", data=fludata)
match1 <- pairmatch(dist1, data=fludata)
match1_summary = summarize.match(fludata, match1)

##1:1 with caliper 
dist1_caliper <- addcaliper(dist1, z=fludata$receive, p=fludata$prop, caliper=0.1)
match1_caliper = pairmatch(dist1_caliper, data=fludata)

match1_caliper_summary = summarize.match(fludata, match1_caliper)
match1_caliper_mean_max = mean_and_max(fludata, match1_caliper)
match1_caliper_dist_mean = round(match1_caliper_mean_max[["mean"]],4)
match1_caliper_dist_max = round(match1_caliper_mean_max[["max"]],4)

match1_caliper_diffs = match1_caliper_mean_max[["all"]] %>% select(id.0, id.1, prop.1, prop.0) %>% mutate(diff = prop.1 - prop.0)

#Plot covariate balance before and after 1:1 matching
png(filename="matching/balance_pre_post_1_1_calip.png")
plot(xBalance(receive ~ assign + age + copd + dm + heartd + race + renal + sex + liverd -1, strata=list(unstrat=NULL, match1_calip=~match1_caliper), data=fludata)) #BEFORE AND AFTER
title('Covariate Balance Before and After 1:1 matching with caliper')
legend(
    "topright",
    legend = c("unstrat", "1:1 matching with caliper"),
    inset = 0.01,
    pch = c(15, 16),
    bg = "white"
)
dev.off()
```

##Generate 1:k dataset 

```{r}
#Generate baseline 1:k dataset
match1_1_k = pairmatch(dist1_caliper, controls = 2, data=fludata)
match1_1_k_summary = summarize.match(fludata, match1_1_k)

#Plot covariate balance before and after 1:k matching
##TO-DO: Replace k with correct number once we set a definitive k. 
#save plot 
png(filename="matching/balance_pre_post_1_k.png")
plot(xBalance(receive ~ assign + age + copd + dm + heartd + race + renal + sex + liverd -1, strata=list(unstrat=NULL, match1_1_k=~match1_1_k), data=fludata)) #BEFORE AND AFTER
title('Covariate Balance Before and After Matching with k = 2')
legend(
    "topright",
    legend = c("unstrat", "1:k matching"),
    inset = 0.01,
    pch = c(15, 16),
    bg = "white"
)
dev.off()

#Generate new dataset forcing balance on COPD
dist1_k_penalty <- addalmostexact(dist1_caliper, z=fludata$z, f=fludata$copd)
match1_1_k_5_exact_copd = pairmatch(dist1_k_penalty, controls = 2, data=fludata)
match1_1_k_5_exact_copd_summary = summarize.match(fludata, match1_1_k_5_exact_copd)

png(filename="matching/balance_pre_post_1_k_copd_exact.png")
plot(xBalance(receive ~ assign + age + copd + dm + heartd + race + renal + sex + liverd -1, strata=list(unstrat=NULL, match1_1_k_5_exact_copd=~match1_1_k_5_exact_copd), data=fludata)) #BEFORE AND AFTER
title('Covariate Balance Before and After 1:k matching with copd exact')
legend(
    "topright",
    legend = c("unstrat", "1:1 matching with copd exact"),
    inset = 0.01,
    pch = c(15, 16),
    bg = "white"
)
dev.off()

```

```{r}
match1_mean_max = mean_and_max(fludata, match1)
match1_dist_mean = round(match1_mean_max[["mean"]],4)
match1_dist_max = round(match1_mean_max[["max"]],4)

match1_diffs = match1_mean_max[["all"]] %>% select(id.0, id.1, prop.1, prop.0) %>% mutate(diff = prop.1 - prop.0)
```

#Estimate treatment effect 

##Estimate ATT for 1:1, 1:1 caliper, 1:k, 1:k with exact COPD

```{r}
att = match1_summary %>% mutate(t_i = outcome.1 - outcome.0)  %>% summarize(att = mean(t_i)) %>% pull(att)
att_caliper = match1_caliper_summary %>% mutate(t_i = outcome.1 - outcome.0)  %>% summarize(att = mean(t_i)) %>% pull(att)
att_1_k = match1_1_k_summary %>% mutate(t_i = outcome.1 - outcome.0)  %>% summarize(att = mean(t_i)) %>% pull(att)
att_1_k_5_exact_copd = match1_1_k_5_exact_copd_summary %>% mutate(t_i = outcome.1 - outcome.0)  %>% summarize(att = mean(t_i)) %>% pull(att)
```

##Compute bias-corrected ATT
```{r}
#Compute different bias-corrected ATTs for different samples
adjustment_1_to_1 = bias_adjusment(data = fludata, match_data = match1_summary)
adjustment_1_to_1_calip = bias_adjusment(data = fludata, match_data = match1_caliper_summary)
adjustment_1_to_k = bias_adjusment(data = fludata, match_data = match1_1_k_summary)
adjustment_1_to_k_5_exact_copd = bias_adjusment(data = fludata, match_data = match1_1_k_5_exact_copd_summary)

#Show unadjusted and adjusted ATT for the different datasets 
adj_att_1_to_1 = adjustment_1_to_1[["adjusted_att"]]
unadj_att_1_to_1 = adjustment_1_to_1[["unadjusted_att"]]
ci_1_to_1 = adjustment_1_to_1[["ci"]]

adj_att_1_to_1_calip = adjustment_1_to_1_calip[["adjusted_att"]]
unadj_att_1_to_1_calip = adjustment_1_to_1_calip[["unadjusted_att"]]
ci_1_to_1_calip = adjustment_1_to_1_calip[["ci"]]

adj_att_1_to_k = adjustment_1_to_k[["adjusted_att"]]
unadj_att_1_to_k = adjustment_1_to_k[["unadjusted_att"]]
ci_1_to_k = adjustment_1_to_k[["ci"]]

adj_att_1_to_k_5_exact_copd = adjustment_1_to_k_5_exact_copd[["adjusted_att"]]
unadj_att_1_to_k_5_exact_copd = adjustment_1_to_k_5_exact_copd[["unadjusted_att"]]
ci_1_to_k_5_exact_copd = adjustment_1_to_k_5_exact_copd[["ci"]]
```

##Implement Rosenbaum's sensitivity analysis

```{r}
#Test whether I am implementing it well on simulated data
# testing = data.frame(outcome.1 = rbinom(100, 1, 0.2), outcome.0 = rbinom(100, 1, 0.1))
# senmv(testing,gamma=3,trim=1)
#Yep, working as expected. If very low treatment effect, high p-value, and viceversa.


match1_to_1_sens = match1_summary %>% select(outcome.1, outcome.0)
senmv(match1_to_1_sens,gamma=3,trim=1)




data.frame(
  col1 = c()
)

```



#2. Just Do OLS 
```{r}
#Estimate the effect via OLS
ols_results = summary(lm(outcome ~ receive, data=fludata))[["coefficients"]]
#Now with controls
ols_results_controls = summary(lm(outcome ~ receive + age + copd + dm + heartd + race + renal + sex + liverd, data=fludata))[["coefficients"]]
```

